---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
html_document: "PA1_template.html"
keep_md: true
---


## Loading and preprocessing the data

```{r load_packages, warning=FALSE}
Sys.setlocale("LC_TIME", "English")
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(lattice)
```

```{r load_data, cache=TRUE}
# Download file
#download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip")
  
# Unzip file
unzip("activity.zip")

# load data
acts <- read.csv("activity.csv")

str(acts)

```

### Process/transform the data

```{r}
# mutate date to the date format

acts <- mutate(acts, dat = ymd(date))
```

## What is mean total number of steps taken per day?

## The mean total number of steps taken per day is 9354.

```{r}
total_steps <- group_by(acts,dat) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE))
summary(total_steps)

```

## What is the average daily activity pattern?

## about 10% of the days appear to be no movement, steps = 0, might be due to monitor faliure(?). Otherwise, most of the days, about 9%, the activity takes 10000 steps in total. 

```{r}
#plot a histogram of total_steps by day
g <- ggplot(total_steps, aes(total_steps))
g + geom_histogram()+ggtitle("daily activity pattern")

```

## Imputing missing values

### using mean of 5-minute interval for the missing value

```{r}
# get the mean of 5-minute interval 
int_avg <- group_by(acts,interval)%>%
  summarise(mean_steps = mean(steps, na.rm = TRUE))
head(int_avg)

#make a copy
acts1 <- acts

#give the missing steps of that interval the mean steps of that interval
for (i in 1:nrow(acts1)){
  t <- acts1$steps[i]
  if (is.na(t)){
    t_int <- acts1$interval[i]
    t_step <- subset(int_avg,interval==t_int)
    t <- t_step$mean_steps[1]
    acts1$steps[i] <- t
  } 
}

head(acts1)
```


## Are there differences in activity patterns between weekdays and weekends?

### ans: the activity peak hour on both weekdays and weekends appears around 8-9am. for the rest of the day, activities are more often in the afternoon and evening (especially around 8pm) on weekends. On weekdays, 5-8am and 6-7pm are quite busy, probably on the way to and back from work.

```{r}
#add a factor as weekends
weekends <- c("Sat", "Sun")
acts1$weekends <- factor((weekdays(acts1$dat,abbreviate = TRUE) %in% weekends),
                        levels=c(FALSE, TRUE), 
                        labels=c('weekdays','weekends'))

head(acts1)
#group the data by inerval and weekends
acts1_st <- group_by(acts1,weekends,interval)%>%
  summarise(avreage_steps = mean(steps))
head(acts1_st)
```

```{r}
#plot the activity by intervals for weekdays and weekends
g <- ggplot(acts1_st, aes(x=interval, y=avreage_steps))
g + geom_line(aes(color=weekends))
```
